---
title: "Daily Exercise 09: Linear Modeling of Ozone Concentration"
author: "Mckenna Cooper"
date: "2025-03-03"
format: html
execute:
  echo: true
---

```{r}
library(tidyverse)
library(visdat)
library(broom)
```

## 1.

```{r}
data(airquality)
?airquality
```

## 2.

```{r}
library(naniar)
vis_miss(airquality)
```

vis_dat was only resulting in errors for me so I used vis_miss instead. This data does need cleaning since Ozone and Solar R. are missing some values.

```{r}
clean_airquality <- na.omit(airquality)
```

## 3.

I choose temperature as my predictor variable because ozone levels often correlate with temperature due to atmospheric conditions.

```{r}
model <- lm(Ozone ~ Temp, data = clean_airquality)
```

## 4.

```{r}
summary(model)
```

The p-value of temp is very small (\<2e-16), meaning temperature significantly affects ozone levels. Due to this, this model seems valid.

## 5.

Here, **R² = 0.488** means \~48.8% of the variance in `Ozone` is explained by `Temp`.

## 6.

```{r}
a <- augment(model, clean_airquality)
```

## 7.

```{r}
library(ggplot2)

ggplot2::ggplot(a, aes(x = Ozone, y = .fitted)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red") + 
  labs(title = "Actual vs. Predicted Ozone Levels",
       subtitle = paste("Correlation:", round(cor(a$Ozone, a$.fitted), 2)),
       x = "Actual Ozone",
       y = "Predicted Ozone")
```
